```{r echo=FALSE}
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
options(stringsAsFactors = FALSE)

transform_counts = function (counts, ...)
    mutate_each_(counts, funs_(lazyeval::lazy_dots(...)), col_data$Sample)

summarize_each = summarise_each
```

We are going to use a small toy data set.

```{r}
experiment = data.frame(
    Gene = LETTERS[1 : 5],
    Length = c(1000, 1000 * (1 : 4)),
    lib1 = c(1000, 1000 * (1 : 4)),
    lib2 = c(1000, 1000 * (1 : 4)),
    lib3 = c(1000, 1000 * (1 : 4)),
    lib4 = c(2000, 1000 * (1 : 4))
)

col_data = data.frame(
    Sample = paste0('lib', 1 : 4),
    Condition = rep(c('control', 'treatment'), each = 2)
)
```

The `col_data` describes the *column data* of our experiment — in other words,
the experimental setup.

---

Here is how the data is transformed into FPKM. All calculations are performed in
log domain to avoid loss of precision in floating point calculations.

```{r}
fpkm = function (counts, transcript_lengths)
    exp(log(counts) - log(transcript_lengths) - log(sum(counts)) + log(1e9))

(fpkm_counts = transform_counts(experiment, fpkm(., Length)))
```

Here are TPM:

```{r}
tpm = function (counts, transcript_lengths) {
    log_by_size = log(counts) - log(transcript_lengths)
    exp(log_by_size - log(sum(exp(log_by_size))) + log(1E6))
}

(tpm_counts = transform_counts(experiment, tpm(., Length)))
```

And finally, library size factors. Unlike the previous methods, this uses data
across samples, and calculates per-sample normalisation factors.

```{r}
size_factors = function (samples) {
    log_samples = log(samples)
    log_means = rowMeans(log_samples)
    summarize_each(log_samples, funs(exp(median(. - log_means))))
}

(sf = size_factors(select(experiment, one_of(col_data$Sample))))
```
