\section{Introduction}

Put non-coding RNAs and pol~III-driven transcription into the context of genomics.

\subsection{Gene expression and regulation}

\subsection{Using high-throughput sequencing for whole-genome analysis}

\subsection{Types of non-coding RNA, their function and expression}

\subsection{Sections in this thesis}

\section{Motivation}

\subsection{The pol~III transcriptome and pol~III-driven transcription}

Expand on the introduction, with a special focus on tRNAs.

\subsection{Using high-throughput sequencing technologies to profile gene expression}

Summary of the state of the art of high-throughput sequencing with a focus on
RNA-seq and ChIP-seq to quantify expression. Explain both the fundamentals of
the experimental side, as well as the bioinformatical toolchain.

\section{Developmental stability of the mRNA–tRNA interface}

Summarise the \emph{Genome Research} paper.

\subsection{tRNA gene expression is actively differentially regulated in mammals}

Focus on the biological question driving the research, and how the paper
contributes to the understanding of the transcriptional variability and
regulation of tRNA genes, and implications of the regulation on protein-coding
genes.

\subsection{Using high-throughput sequencing data to analyse tRNA gene expression}

Explain the particular computational and statistical challenges encountered
when analysing the experimental data. This includes a discussion of the
decision to profile tRNA gene expression using pol~III ChIP-seq rather than
total RNA-seq, as well as customisations to the general computational sequence
analysis pipeline, which were necessary to tackle the data.

Lastly, this section also includes a description of the specific analysis steps
performed in the interpretation of the count data (expanding on the “Methods”
section of the paper) as well as their limitations.

\subsection{A technical overview of failed approaches to the analysis}

\section{An analysis of repeat elements and their expression}

Depends on results; first section should focus on unique challenges of this
analysis due to the bad mappability of repeat analysis, and approaches chosen
to mitigate this (and how that impacts interpretation).

\subsection{Technical challenges of repeat expression quantification}

\subsection{SINE expression}

\section{Expression profiles of long non-coding RNAs through development}

\section{Summary and concluding remarks}
